Fixing Vulnerabilities Hinders Maintainability

### 2020-01-31 paper/icpc20.tex
new version

example of a security refactoring of a known vulnerability found in
OpenSSL\footnote{\label{openssl}OpenSSL is a toolkit that

@@

%%%%% NEEDS REVIEW
As an example, consider the refactoring of the Datagram Transport Layer Security
(DTLS) protocol implementation in OpenSSL\footnoteref{openssl} to address a vulnerability where an attacker could cause a Denial-of-Service
(DoS) attack by crafting DTLS handshake messages to trigger memory allocations
corresponding to large length values. This vulnerability is listed at the Common
Vulnerabilities and Exposures dictionary as CVE-$2014$-$3506$\footnote{CVE-$2014$-$3506$
details available at \url{http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2014-3506}
(Accessed on \today{})}. It is amongst the vulnerabilities studied in our
evaluation. The snippet, in Listing~\ref{lst:vuln}, presents the changes performed on the
\emph{ssl/d1\_both.c} file\footnote{CVE-$2014$-$3506$ fix available  at
\url{https://github.com/openssl/openssl/commit/1250f12613b61758675848f6600ebd914ccd7636}
(Accessed on \today{})} by the OpenSSL developers to patch the vulnerability. Every SSL/TLS connection begins with a handshake which is responsible by the
negotiation between the two parties. DTLS provides a mechanism for fragmenting
a handshake message over a number of packets, each of which can be transmitted
separately over the internet from one party to another. The problem in
CVE-$2014$-$3506$ is that the size of the handshake message collected by the
receiver is not verified against the maximum size allowed by DTLS. It is then
allocated to a new fragment (cf. Listing~\ref{lst:vuln}, line~36). Thus, if an
attacker has access to one of these packages, he can change the
handshake message size and trigger unnecessary memory allocations of large
length which can lead to memory exhaustion that could end in a DoS attack.

\medskip
\setcounter{lstannotation}{0}
\begin{lstlisting}[style={JavaStyle}, caption={Fix provided by Netty developers to the
\\CVE-$2014$-$3488$ vulnerability},label={lst:vuln}]
// [snip]
   for (;;) {
+       final int outAppBufSize = engine.getSession().getApplicationBufferSize();
+       final ByteBuffer outAppBuf;
+       if (nioOutAppBuf.capacity() < outAppBufSize) {
+              outAppBuf = ByteBuffer.allocate(outAppBufSize);
+       } else {
+              outAppBuf = nioOutAppBuf;
+       }

         try {
-               result = engine.unwrap(nioInNetBuf, nioOutAppBuf);
-               result = engine.unwrap(nioInNetBuf, outAppBuf);
// [snip]
                 break;
         } finally {
-               nioOutAppBuf.flip();
+              outAppBuf.flip();

                nettyInNetBuf.readerIndex(nettyInNetBufStartOffset + nioInNetBuf.position() - nioInNetBufStartOffset);

-               if (nioOutAppBuf.hasRemaining()) {
+              if (outAppBuf.hasRemaining()) {
                       if (nettyOutAppBuf == null) {
                              ChannelBufferFactory factory = ctx.getChannel().getConfig().getBufferFactory();
                              nettyOutAppBuf = factory.getBuffer(initialNettyOutAppBufCapacity);
                       }
-                    nettyOutAppBuf.writeBytes(nioOutAppBuf);
+                    nettyOutAppBuf.writeBytes(outAppBuf);
                 }
-              nioOutAppBuf.clear();
+              outAppBuf.clear();
          }
     }

%\medskip
%\setcounter{lstannotation}{0}
%\begin{lstlisting}[style={JavaStyle}, caption={Fix provided by OpenSSL developers to the
%\\CVE-$2014$-$3506$ vulnerability},label={lst:vuln}]
%+ static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
%+  unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
%+  if(max_len < (unsigned long)s->max_cert_list)
%+     return s->max_cert_list;
%+  return max_len;
%+ }
%
%  static int dtls1_reassemble_fragment(
%     SSL *s, struct hm_header_st* msg_hdr, int *ok){
%
%// [snip]
%
%- unsigned long frag_len = msg_hdr->frag_len, max_len;
%- if((msg_hdr->frag_off+frag_len) > msg_hdr->msg_len)
%-    goto err;
%
%- if(DTLS1_HM_HEADER_LENGTH +
%-    	SSL3_RT_MAX_ENCRYPTED_LENGTH < s->max_cert_list)
%-    max_len = s->max_cert_list;
%- else
%-    max_len = DTLS1_HM_HEADER_LENGTH +
%-   		SSL3_RT_MAX_ENCRYPTED_LENGTH;
%
%+ unsigned long frag_len = msg_hdr->frag_len;
%
%- if((msg_hdr->frag_off+frag_len) > max_len)
%+ if((msg_hdr->frag_off+frag_len) > msg_hdr->msg_len ||
%+ msg_hdr->msg_len > dtls1_max_handshake_message_len(s))
%     goto err; /*!\annotation{lst:func5}!*/
%
%  memset(seq64be,0,sizeof(seq64be));
%  seq64be[6] = (unsigned char) (msg_hdr->seq>>8);
%  seq64be[7] = (unsigned char) msg_hdr->seq;
%  item = pqueue_find(s->d1->buffered_messages, seq64be);
%  if(item == NULL){
%     frag = dtls1_hm_fragment_new(msg_hdr->msg_len, 1);
%
%// [snip]
%
% static int dtls1_process_out_of_seq_message(
%     SSL *s, struct hm_header_st* msg_hdr, int *ok){
%
%// [snip]
%
%+ if(frag_len > dtls1_max_handshake_message_len(s))
%+    goto err; /*!\annotation{lst:func4}!*/
%
%  frag = dtls1_hm_fragment_new(frag_len, 0);
%
%\end{lstlisting}

The code changes performed to fix the vulnerability are:

>>>>

example of a security refactoring of a known vulnerability found in the
PHP interpreter\footnote{\label{openssl}OpenSSL is a toolkit that

@@

%%%%% NEEDS REVIEW
As an example, consider the patch of the state-machine implementation in OpenSSL\footnoteref{openssl} to address a vulnerability where an attacker could cause a Denial-of-Service
(DoS) attack by crafting TLS messages to trigger memory allocations
corresponding to large length values. This vulnerability is listed at the Common
Vulnerabilities and Exposures dictionary as CVE-$2014$-$3506$\footnote{CVE-$2014$-$3506$
details available at \url{http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2014-3506}
(Accessed on \today{})}. It is amongst the vulnerabilities studied in our
evaluation. The snippet, in Listing~\ref{lst:vuln}, presents the changes performed on the
\emph{ssl/d1\_both.c} file\footnote{CVE-$2014$-$3506$ fix available  at
\url{https://github.com/openssl/openssl/commit/1250f12613b61758675848f6600ebd914ccd7636}
(Accessed on \today{})} by the OpenSSL developers to patch the vulnerability. Every SSL/TLS connection begins with a handshake which is responsible by the
negotiation between the two parties. DTLS provides a mechanism for fragmenting
a handshake message over a number of packets, each of which can be transmitted
separately over the internet from one party to another. The problem in
CVE-$2014$-$3506$ is that the size of the handshake message collected by the
receiver is not verified against the maximum size allowed by DTLS. It is then
allocated to a new fragment (cf. Listing~\ref{lst:vuln}, line~36). Thus, if an
attacker has access to one of these packages, he can change the
handshake message size and trigger unnecessary memory allocations of large
length which can lead to memory exhaustion that could end in a DoS attack.

\medskip
\setcounter{lstannotation}{0}
\begin{lstlisting}[style={CStyle}, caption={Fix provided by OpenSSL developers to the
\\CVE-2016-6307 vulnerability for file ssl/statem/statem.c},label={lst:vuln}]
 static SUB_STATE_RETURN read_state_machine(SSL *s){ /*!\annotation{lst:func1}!*/
 // [snip]
+   if (!SSL_IS_DTLS(s) && s->s3->tmp.message_size > 0
+        && !BUF_MEM_grow_clean(s->init_buf, 
+                                (int)s->s3->tmp.message_size
+                                + SSL3_HM_HEADER_LENGTH)) {
+         ssl3_send_alert(s, SSL3_AL_FATAL, SSL_AD_INTERNAL_ERROR);
+         SSLerr(SSL_F_TLS_GET_MESSAGE_HEADER, ERR_R_BUF_LIB);
+          return SUB_STATE_ERROR;
 // [snip]
 }
\end{lstlisting}
%
%\medskip
%\setcounter{lstannotation}{0}
%\begin{lstlisting}[style={CStyle}, caption={Fix provided by OpenSSL developers to the
%\\CVE-2016-6307 vulnerability for file ssl/statem/statem_lib.c },label={lst:vuln}]
% int tls_get_message_header(SSL *s, int *mt){
% // [snip]
%+   if (!SSL_IS_DTLS(s) && s->s3->tmp.message_size > 0
%+        && !BUF_MEM_grow_clean(s->init_buf, 
%+                                (int)s->s3->tmp.message_size
%+                                + SSL3_HM_HEADER_LENGTH)) {
%+         ssl3_send_alert(s, SSL3_AL_FATAL, SSL_AD_INTERNAL_ERROR);
%+         SSLerr(SSL_F_TLS_GET_MESSAGE_HEADER, ERR_R_BUF_LIB);
%+          return SUB_STATE_ERROR;
% // [snip]
% }
%\end{lstlisting}

%The code changes performed to fix the vulnerability are:
